### 2020-03-09 document/manuscript.tex
Travis build: 67

\section{Towards Open and Reproducible Genomic Research: Lessons from
OpenScienceKE}\label{towards-open-and-reproducible-genomic-research-lessons-from-openscienceke}

\subsection*{Abstract}\label{abstract}
\addcontentsline{toc}{subsection}{Abstract}

Open, collaborative and reproducible research -- Open Science -- has a
great potential for advancing science. However, the training in our
local universities does not equip students with the tools to practice
open science. However, to work in the open and collaborate, your
collaborators should be equipped to use the tools that you use. The main
barrier to working open, therefore, is the lack of awareness of the
collaboration tools and the skills required to utilize these tools.
Therefore, to fill the gap through an open science community, funded by
a Mozilla Mini-grant -- OpenScienceKE, we are promoting open science
among bioinformatics students and researchers in the Nairobi area by
training using this model: sensitize, train, hack and collaborate. This
model first sensitizes on open science practices through seminars,
trains on open science tools through workshops, facilitates hands-on
application of the tools through hackathons, and finally fosters a
community of open science enthusiasts through meetups.

OpenScienceKE sought to address the following problems: the lack of
awareness of open science practices and tools within the Bioinformatics
community in Kenya; the poor adoption of open science practices in
Bioinformatics; and the absence of research to establish the state of
affairs in adopting open science in Kenya. From the OpenScienceKE
hackathon, we managed to create an open resource that the students could
use to figure out where they can cost-effectively publish open access.
In addition, through literature search and data mining, we observed a
growing interest in open science practices in Kenya but the lack of
awareness and skills hinder the adoption. The use of preprints for
research dissemination haven't caught up in Kenya; out of the 20,069
papers downloaded from bioRXiv, only 18 have Kenyan authors, a majority
of which are as a result of international (16) collaborations. We also
observed a lack of incentives and policy in academic and research
institutions to support open science. The fear of being scooped and the
competitive spirit within the scientific community are also major
barriers to working in the open.

The first iteration of the model which focused on academic institutions
set the foundation for next phase: promote the open and reproducible
science in research institutions. This model provides the framework for
the adoption of open science practices within the institution and others
in the future. As genomic research data generated in Africa grows, there
is a need for the adoption of open science practices in data storage,
reproducible pipelines and collaborative research. We propose this
approach, which develops the necessary infrastructure within research
institutions, and builds human capacity through the model: sensitize,
train, hack and collaborate. Promotion of open science in Africa
recognizes the future direction of research and OpenScienceKE is growing
the culture and practice in the research active region.

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

Our Core message is to present a clear outlook on the adoption of Open
Science practice by African-based scientists, starting with Kenya.

Add an introduction of the Open Science. To cite an article, use
\cite{Schlegel2016}. All the bibliographies should be added to
\texttt{library.bib} in the BibTeX format. See the example in
\texttt{library.bib}.

Since this will more or less be like a review article, we will need to
identify the various subsections based on the topics we need to cover in
the review. See
\href{http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005619}{this
article} for some tips \cite{Mensh2017}.

\subsection*{Status of Open Science in Kenya: Literature
search}\label{status-of-open-science-in-kenya-literature-search}
\addcontentsline{toc}{subsection}{Status of Open Science in Kenya:
Literature search}

The report from the above team will be useful for writing the
introduction, as well as providing materials to be used in the
discussion. In fact, we can use their resource to weave the whole paper
together.

In this section, and the introduction, we will conduct a review of the
status of open science in the country. - What kind of resources are
available to support open science - Are there policies or incentives for
open science practices? - What kind of training activities have been
conducted to promote and train students and researchers on open science
tools and practices? - etc

\section*{DataMining Section}\label{datamining-section}
\addcontentsline{toc}{section}{DataMining Section}

This will be a data analysis section. The title of this section will
depend on the results of your analysis. For example, if Kenyan
researchers are not publishing open access, we will need to understand
why that is the case. The solution may lie in the cost of publishing,
and that is how the resource created by Open Access options team is
useful.

We address questions like: - What is the publishing trend by Kenyan
researchers - Are they publishing open access, and how has this changed
over the years? - Are Kenyan researchers embracing pre-prints (BioRXiv,
AriRXiv, ResearchGate, F1000Research). Who is driving the adoption of
pre-prints? Local researchers or foreign collaborators? - What are the
collaboration trends? Are Kenyan researcher collaborating locally or
internationally?

\section*{But Publishing Open Access is
Expensive}\label{but-publishing-open-access-is-expensive}
\addcontentsline{toc}{section}{But Publishing Open Access is Expensive}

The article processing charge is the main barrier to publishing open
access, in addition to the obsession with impact factors. However, for
early career scientists and students, especially in developing
countries, most publishers offer waivers and subsidies but few are
aware. In this section, we explore some of the avenues to publishing
open access at low cost.

To address this problem, we created a resource that can guide ECR and
students on where they publish open access, and at low cost. We also
provide information on how they can still be open when they publsih in
paywalled journals, eg via the green route.

\section*{Figures}\label{figures}
\addcontentsline{toc}{section}{Figures}

You can add the figures as follows:

\begin{figure}[htbp]
\centering
%% \includegraphics{./figure01.pdf}
\caption{Figure 1}
\end{figure}

And you can have it referenced as a figure

\begin{quote}
\textbf{Box 1} To highlight of defining some key concepts in Open
science without disrupting the flow of the articles, you can use a quote
format.
\end{quote}

\section*{Discussion}\label{discussion}
\addcontentsline{toc}{section}{Discussion}

What do the results mean? How does your results fit to the current
literature? How do they compare to other similar studies?

\section*{Conclusions}\label{conclusions}
\addcontentsline{toc}{section}{Conclusions}

What is the take-home message from this article? What are the
recommendations? - The need for a framework to guide the adoption of
open science practices. Take note of the bariers and provide
recommendations. - The need for low-cost publishing - The need for
policies on open science that can be implemented by various
insituitions. For example, we can provide a template that can be adopted
by most instituitions

\section{Acknowledgement}\label{acknowledgement}

We acknowledge the support from \ldots{} and the contribution from
Mozilla Science Lab and KENET.

Use this section to acknowledge funding and resource contributions to
the project.